Component response rate variation underlies the stability of complex systems

### 2020-03-15 notebook/SI.Rmd
Report the SWN results and add the network figures to the SI

Structured network construction {#structured}
================================================================================

Example of a small-world network

```{r, echo = FALSE}
source("../R/small_world.R");
par(mfrow = c(1, 2), mar = c(5, 5, 1, 1));
swn <- create_swn(S = 72, K = 12, beta = 0.01);
visualise_network(swn);
swn <- create_swn(S = 72, K = 12, beta = 0.1);
visualise_network(swn);
```

>>>>

Stability of structured networks {#structured}
================================================================================

I tested the stability of one million random, small-world, scale-free, and cascade food web networks. Each of these networks is structured differently. In the main text, the random networks and cascade food webs that I built were saturated ($C = 1$), meaning that every component was connected to, and interacted with, every other component.

\vspace{2mm}

```{r, echo = FALSE, fig.height = 6, fig.width = 6, fig.align = "center"}
A0_dat <- rnorm(n = 20 * 20, mean = 2, sd = 0);
A0     <- matrix(data = A0_dat, nrow = 20, ncol = 20);
visualise_network(A0);
```

\vspace{2mm}

Small-world networks, in contrast, are not saturated. They are instead defined by components that interact mostly with other closely neighbouring components, but have a proportion of interactions ($\beta$) that are instead between non-neighbours. Two small-world networks are shown below.

\vspace{2mm}

```{r, echo = FALSE, fig.align = "center"}
source("../R/small_world.R");
par(mfrow = c(1, 2), mar = c(5, 5, 1, 1));
swn <- create_swn(S = 72, K = 12, beta = 0.01);
visualise_network(swn);
swn <- create_swn(S = 72, K = 12, beta = 0.1);
visualise_network(swn);
```

\vspace{2mm}

The small-world network on the left shows a system in which $\beta = 0.01$, while the small-world network on the right shows one in which $\beta = 0.1$. At the extremes of $\beta = 0$ and $\beta = 1$, networks are regular and random, respectively. The table below shows how $\sigma^{2}_\gamma$ affects stability in small world networks across different values of $S$ and $\beta$.

```{r, echo = FALSE, fig.align = "center"}
swn <- read.csv(file = "sim_results/swn/old/swn.csv"); # Change when new results are in
use <- c(1:6, 19:21);
kable(swn[,use]);
```

In the above, the mean $C$, complexity of $\mathbf{A}$, and complexity of $\mathbf{M}$ are also shown.

### 2020-03-15 notebook/SI.tex
Report the SWN results and add the network figures to the SI

\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}

@@

\setlength{\droptitle}{-2em}

  \title{Component response rate variation underlies the stability of complex
systems}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \subtitle{Supplemental Information}
  \author{A. Bradley Duthie}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{Biological and Environmental Sciences, University of Stirling, Stirling,
UK, FK9 4LA
\href{mailto:alexander.duthie@stir.ac.uk}{\nolinkurl{alexander.duthie@stir.ac.uk}}}

@@

  \begin{itemize}
  \tightlist
  \item
    \protect\hyperlink{sigma3}{\(\sigma\) = 0.3}
  \item
    \protect\hyperlink{sigma4}{\(\sigma\) = 0.4}
  \item
    \protect\hyperlink{sigma5}{\(\sigma\) = 0.5}
  \item
    \protect\hyperlink{sigma6}{\(\sigma\) = 0.6}
  \end{itemize}
\item
  \protect\hyperlink{gam_dist}{Sensitivity of distribution of
  \(\gamma\)}
\item
  \protect\hyperlink{Feasibility}{Feasibility of complex systems}
\item
  \protect\hyperlink{ga}{Stability given targeted manipulation of
  \(\gamma\) (genetic algorithm)}
\item
  \protect\hyperlink{Gibbs}{Consistency with Gibbs et al. (2018)}
\item
  \protect\hyperlink{rhos}{Effects of matrix correlations}
\end{itemize}

@@

@@

\includegraphics{SI_files/figure-latex/unnamed-chunk-24-1.pdf}

@@

{\centering \includegraphics{SI_files/figure-latex/unnamed-chunk-28-1}

@@

\includegraphics{SI_files/figure-latex/unnamed-chunk-32-1.pdf}

@@

\includegraphics{SI_files/figure-latex/unnamed-chunk-33-1.pdf}

@@

@@

\includegraphics{SI_files/figure-latex/unnamed-chunk-38-1.pdf}

\section{Effects of matrix
correlations}\label{effects-of-matrix-correlations}

>>>>

\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\ImportTok}[1]{{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{{#1}}}}
\newcommand{\BuiltInTok}[1]{{#1}}
\newcommand{\ExtensionTok}[1]{{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{{#1}}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\NormalTok}[1]{{#1}}

@@

\setlength{\droptitle}{-2em}
  \title{Component response rate variation underlies the stability of complex
systems}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
\subtitle{Supplemental Information}
  \author{A. Bradley Duthie}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \predate{\centering\large\emph}
  \postdate{\par}
  \date{Biological and Environmental Sciences, University of Stirling, Stirling,
UK, FK9 4LA
\href{mailto:alexander.duthie@stir.ac.uk}{\nolinkurl{alexander.duthie@stir.ac.uk}}}

@@

  \begin{itemize}
  \tightlist
  \item
    \protect\hyperlink{sigma3}{\(\sigma\) = 0.3}
  \item
    \protect\hyperlink{sigma4}{\(\sigma\) = 0.4}
  \item
    \protect\hyperlink{sigma5}{\(\sigma\) = 0.5}
  \item
    \protect\hyperlink{sigma6}{\(\sigma\) = 0.6}
  \end{itemize}
\item
  \protect\hyperlink{gam_dist}{Sensitivity of distribution of
  \(\gamma\)}
\item
  \protect\hyperlink{structured}{Stability of structured networks}
\item
  \protect\hyperlink{Feasibility}{Feasibility of complex systems}
\item
  \protect\hyperlink{ga}{Stability given targeted manipulation of
  \(\gamma\) (genetic algorithm)}
\item
  \protect\hyperlink{Gibbs}{Consistency with Gibbs et al. (2018)}
\end{itemize}

@@

\hypertarget{structured}{\section{Stability of structured
networks}\label{structured}}

I tested the stability of one million random, small-world, scale-free,
and cascade food web networks. Each of these networks is structured
differently. In the main text, the random networks and cascade food webs
that I built were saturated (\(C = 1\)), meaning that every component
was connected to, and interacted with, every other component.

\vspace{2mm}

\begin{center}\includegraphics{SI_files/figure-latex/unnamed-chunk-23-1} \end{center}

\vspace{2mm}

Small-world networks, in contrast, are not saturated. They are instead
defined by components that interact mostly with other closely
neighbouring components, but have a proportion of interactions
(\(\beta\)) that are instead between non-neighbours. Two small-world
networks are shown below.

\vspace{2mm}

\begin{center}\includegraphics{SI_files/figure-latex/unnamed-chunk-24-1} \end{center}

\vspace{2mm}

The small-world network on the left shows a system in which
\(\beta = 0.01\), while the small-world network on the right shows one
in which \(\beta = 0.1\). At the extremes of \(\beta = 0\) and
\(\beta = 1\), networks are regular and random, respectively. The table
below shows how \(\sigma^{2}_\gamma\) affects stability in small world
networks across different values of \(S\) and \(\beta\).

\begin{longtable}[]{@{}rrrrrrrrr@{}}
\toprule
beta & S & A0\_unstable & A0\_stable & A1\_unstable & A1\_stable &
complex\_A0 & complex\_A1 & C\tabularnewline
\midrule
\endhead
0.00 & 24 & 17442 & 982558 & 17489 & 982511 & 0.1695021 & 0.1941223 &
0.1304348\tabularnewline
0.00 & 48 & 257013 & 742987 & 259517 & 740483 & 0.2355489 & 0.2711380 &
0.1063830\tabularnewline
0.00 & 72 & 715587 & 284413 & 723082 & 276918 & 0.2866561 & 0.3304009 &
0.0985915\tabularnewline
0.00 & 96 & 961817 & 38183 & 963276 & 36724 & 0.3299115 & 0.3804889 &
0.0947368\tabularnewline
0.00 & 120 & 998976 & 1024 & 998812 & 1188 & 0.3681518 & 0.4246688 &
0.0924370\tabularnewline
0.00 & 144 & 999998 & 2 & 999991 & 9 & 0.4027471 & 0.4646754 &
0.0909091\tabularnewline
0.00 & 168 & 1000000 & 0 & 1000000 & 0 & 0.4345714 & 0.5014674 &
0.0898204\tabularnewline
0.01 & 24 & 17386 & 982614 & 17433 & 982567 & 0.1695343 & 0.1941452 &
0.1304321\tabularnewline
0.01 & 48 & 255369 & 744631 & 257956 & 742044 & 0.2355569 & 0.2711250 &
0.1063798\tabularnewline
0.01 & 72 & 708407 & 291593 & 716019 & 283981 & 0.2866423 & 0.3303843 &
0.0985884\tabularnewline
0.01 & 96 & 960591 & 39409 & 961685 & 38315 & 0.3299322 & 0.3804641 &
0.0947337\tabularnewline
0.01 & 120 & 998991 & 1009 & 998813 & 1187 & 0.3681428 & 0.4246683 &
0.0924339\tabularnewline
0.01 & 144 & 999997 & 3 & 999993 & 7 & 0.4027696 & 0.4646907 &
0.0909060\tabularnewline
0.01 & 168 & 1000000 & 0 & 1000000 & 0 & 0.4345881 & 0.5014852 &
0.0898172\tabularnewline
0.10 & 24 & 20408 & 979592 & 20478 & 979522 & 0.1692658 & 0.1937917 &
0.1302962\tabularnewline
0.10 & 48 & 238101 & 761899 & 240787 & 759213 & 0.2351005 & 0.2705892 &
0.1062307\tabularnewline
0.10 & 72 & 680138 & 319862 & 685524 & 314476 & 0.2861156 & 0.3297663 &
0.0984350\tabularnewline
0.10 & 96 & 961691 & 38309 & 959740 & 40260 & 0.3293218 & 0.3797666 &
0.0945786\tabularnewline
0.10 & 120 & 999544 & 456 & 999282 & 718 & 0.3674529 & 0.4238479 &
0.0922777\tabularnewline
0.10 & 144 & 999999 & 1 & 999999 & 1 & 0.4019505 & 0.4637655 &
0.0907492\tabularnewline
0.10 & 168 & 1000000 & 0 & 1000000 & 0 & 0.4337704 & 0.5005276 &
0.0896598\tabularnewline
0.25 & 24 & 23923 & 976077 & 24031 & 975969 & 0.1680052 & 0.1922420 &
0.1296696\tabularnewline
0.25 & 48 & 228389 & 771611 & 231312 & 768688 & 0.2331850 & 0.2683721 &
0.1055242\tabularnewline
0.25 & 72 & 666374 & 333626 & 668472 & 331528 & 0.2836431 & 0.3268807 &
0.0977062\tabularnewline
0.25 & 96 & 966462 & 33538 & 961616 & 38384 & 0.3263952 & 0.3763783 &
0.0938386\tabularnewline
0.25 & 120 & 999759 & 241 & 999510 & 490 & 0.3642255 & 0.4201168 &
0.0915315\tabularnewline
0.25 & 144 & 1000000 & 0 & 999999 & 1 & 0.3983171 & 0.4595509 &
0.0899989\tabularnewline
0.25 & 168 & 1000000 & 0 & 1000000 & 0 & 0.4298720 & 0.4960198 &
0.0889060\tabularnewline
1.00 & 24 & 26159 & 973841 & 26309 & 973691 & 0.1588024 & 0.1815081 &
0.1249673\tabularnewline
1.00 & 48 & 211700 & 788300 & 214607 & 785393 & 0.2154544 & 0.2478696 &
0.0991369\tabularnewline
1.00 & 72 & 613966 & 386034 & 615862 & 384138 & 0.2602685 & 0.2998802 &
0.0908174\tabularnewline
1.00 & 96 & 943949 & 56051 & 936840 & 63160 & 0.2984622 & 0.3441159 &
0.0867075\tabularnewline
1.00 & 120 & 999134 & 866 & 998351 & 1649 & 0.3323169 & 0.3833061 &
0.0842580\tabularnewline
1.00 & 144 & 1000000 & 0 & 999997 & 3 & 0.3630114 & 0.4187917 &
0.0826318\tabularnewline
1.00 & 168 & 1000000 & 0 & 1000000 & 0 & 0.3913277 & 0.4515183 &
0.0814745\tabularnewline
\bottomrule
\end{longtable}

In the above, the mean \(C\), complexity of \(\mathbf{A}\), and
complexity of \(\mathbf{M}\) are also shown.

Example of a scale-free network

@@

\includegraphics{SI_files/figure-latex/unnamed-chunk-28-1.pdf}

@@

{\centering \includegraphics{SI_files/figure-latex/unnamed-chunk-32-1}

@@

@@

@@

\includegraphics{SI_files/figure-latex/unnamed-chunk-41-1.pdf}

@@

\includegraphics{SI_files/figure-latex/unnamed-chunk-42-1.pdf}
